EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-05-14 draft.tex
Comments on general structure of the paper, as well as some more comments

\title{EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms}

@@

% Maybe these three sections should be merged; they are all experimental set up and the two next ones start with "experimental"
% (and he next one too) - JJ
% done - Mario
% But also change title? - JJ
\section{Container-based application} 
\label{docker}

We have mentioned several advantages of a cloud-native application for running
multi-population based algorithms, but there are in our experience several
practical issues of this approach. In general, cloud platforms have free-tiers
for their services, and many offer academic discounts to students. But once we
pass to the pay-as-you-go tier, computing resources need to be
managed, and, where possible, its cost minimized.  First,
in many cases,  a credit-card is needed to guarantee the payment of the
resources consumed and someone responsible for assigning quotas and, more
importantly, special keys for using the cloud's API.  If these keys are
compromised or shared by their holders,  additional costs or administrative
procedures are in order. For those cases when a prototype platform is needed,
for instance, students developing a new algorithm variant or experimenting with
a published work, a container-based platform can be more practical. Container
technology enables researchers the local deployment of all the infrastructure
needed to run the experiments. Later, if needed, the same system could be
deployed to the cloud.

% The main advantage of container deployments over virtual machine
% instances is that they add much less overhead to communication between
% them, and they are much more economical to deploy. Scaling and
% monitoring can be done much more easily through cloud orchestration
% systems such as Kubernetes; these are the reasons why we chose it for
% this work.
% I don't see this line of argument. You can do exactly the same with
% virtual machines instead of containers. Also, you can test a virtual
% infrastructure locally and then only when it's needed deploy it to
% the cloud. If you want to give good reasons for using containers
% instead of virtual machines, there are quite a lot non-economic ones
% - JJ
% OK, I will add an issue - Mario
% Did you finally? - JJrehe
% I did not, will add and assign it - M
There are many advantages of container-based architectures over one
based on virtual machines. The main ones are architectural: it makes
the design of an application gravitate towards a set of decoupled
tasks, that use reactive, event based programming, to interact with
the rest of the application; containers hold micro-services that are
much easier to test, integrate and deploy. It also simplifies
the interaction APIs. Of course, that makes them more economical in a
pay-per-use cloud provider, but that is not the main reason why we
have chosen them in this paper; reproducibility of experiments and a
clear and idempotent description of the architecture through docker
compose are the main reasons.

In this section, we propose a design of a reactive container-based application
for executing multi-population based optimization experiments.  For the design,
we followed the design patterns highlighted in the previous section, and again
we can go back to Figure \ref{fig:kafkEO} and use it as a guide for explaining
the main components and their interactions.

@@

Figure \ref{fig:experiment_flow} shows the general workflow to run experiments
in Evo\-Swarm, and the three roles users can play. The first step is to develop the
search strategy algorithms that are going to be used by the multi-population
algorithm; algorithm developers do this. Then administrators are responsible for
the installation and configuration of the application. Once the application is
running, users can start one or several experiments, monitor the execution, and
analyze the results. Depending on the type of experiment executed, users could
have additional needs, in this case there are additional scripts for
processing the results of experiments to be used with the COCO framework. In the
next sections, we give more details about these roles and their workflows.

@@

@@

\subsection{Experiment Execution}
\label{sec:exeperimentalDefinition}

@@

\subsection{Using the COCO Framework}

@@

%

@@

We aimed to learn if the proposed solution can efficiently
improve the scalability and performance of population-based optimization
algorithms. % If you want to prove it *improves*, you need to provide a baseline. 
            %What will be your baseline? You need to specify anyway 
            %in which way it really improves - JJ
            % The baseline is a single population GA or PSO? - M
Hence, in the following sections, we verify the following:

@@

\begin{itemize}
\item Can we improve the execution time of the algorithm by adding populations
 and serverless functions to the system and, in particular, what is its effect on the scalability of the system?

\item Does having a multi-population enabled platform, with heterogeneous populations and
the support for mixing search strategies,  increases the performance of the
search by needing fewer function evaluations than a homogeneous setting? % But in this case the answer is no, right? - JJ
                                                                         % Is yes, but is in the second experiment - M
\item Can EvoSwarm as a cloud-native, container-based application,  be deployed
locally or in the cloud by just specifying the configuration of container images
and resources to an orchestration service?
\end{itemize}

@@

\section{Measuring Scalability}
\label{sec:exp1}
% Experiment 1
As we mentioned before, reactive systems can scale by adding additional copies
of serverless functions. In our case, we can start additional worker containers
to have the same effect. Other authors, like Salza et al. \cite{salza2019speed},   % add
                                % reference - JJ
                                % done - Mario
have used a
similar architecture, which implies that worker nodes need to have at least a certain level
of complexity, in terms of execution time,  to effectively scale on multiple
nodes, tending to linear scalability. In that paper, Salza et al. tested scalability
by simulating the work of nodes by using a sleep function and messages of different sizes.
In a population-based algorithm like EvoSwarm, several parameters can increase or 
decrease the workload of workers:

@@

To evaluate the effect of these parameters on the scalability of the system, we
propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
BBOB testbed. This function has been used because it is computationally demanding,
and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
\cite{nicolau2009application} to solve with the required FEs. As we
are only comparing in terms of execution time, the results from a single function
can be better understood, since there are fewer factors involved.
Following the procedure described in the previous section, we executed 
two sets of experiments with five and ten initial populations,
repeating each experiment 30 times using  1, 2, 4, and 8 workers.  We kept the same 
maximum number of FEs, changing the relevant parameters for this. 
See Tables \ref{tab:params:10} and \ref{tab:params:5} 
for the complete list. 
%
\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{time_diff}
  \caption{Boxplot of the median time required for the execution of 30 
  instances of function $f_4$, in seconds, different number of workers and populations sizes.  }
  % Which function? - JJ
  \label{fig:spworker}
\end{figure}
%
\begin{table}[]
  \caption{The median of the time required to finish an instance of the function $f_4$, 30 runs
  with different number of workers and populations}
  \label{tab:time}
  \vspace{0.25cm}
    \centering
    \begin{tabular}{ccrr}
    \hline
              &         & 5 Populations    & 10 Populations \\
     Dimension& Workers & Median of rate & Median of rate  \\
    \hline
          2   & 1w      & 2.96   & 3.15   \\
              & 2w      & 1.84   & 1.72   \\
              & 4w      & 1.22   & \textbf{0.96}   \\
              & 8w      & 1.05   & \textbf{0.59}   \\
    \hline
          3   & 1w      & 6.53   & \textbf{7.76}   \\
              & 2w      & 3.64   & 3.83   \\
              & 4w      & 2.33   & 2.26   \\
              & 8w      & 2.17   &  \textbf{1.36}   \\
    \hline
          5   & 1w      & \textbf{12.33}  & 14.19  \\
              & 2w      & 6.63   & \textbf{7.41}   \\
              & 4w      & 4.25   & 4.34   \\
              & 8w      & 3.67   &  \textbf{2.41}   \\
    \hline
          10  & 1w      & \textbf{31.50}  & 38.30  \\
              & 2w      & \textbf{16.07}  & 19.31  \\
              & 4w      &  9.99   & 10.35  \\
              & 8w      & 8.97   &  \textbf{5.92}  \\
    \hline
          20  & 1w      &  116.35 & 124.51 \\
              & 2w      & \textbf{51.18}  & 63.68  \\
              & 4w      & 32.26  & 33.07  \\
              & 8w      & 28.64  & \textbf{18.25}  \\
    \hline
          40  & 1w      & \textbf{369.89} & 448.89 \\
              & 2w      & \textbf{198.43} & 234.50 \\
              & 4w      & 122.39 &  120.62 \\
              & 8w      & 122.31 & \textbf{63.35} \\
    \hline
  \end{tabular}
\end{table}

@@

% I think this is not the case? 
% Another additional issue is related to the algorithm itself. Many more
% workers than populations will have a wider range of states of
% evolution in the experiment; populations that will be mixed will have
% very different fitness states, which means that a part of them will
% probably have been not so useful, following the intermediate
% disturbance hypothesis. When the number of workers is increased, the
% difference between the state of evolution of populations will be
% intermediate, leading to better results; eventually, when there are as
% many workers as populations, the state will be quite similar, which is
% shown with a smaller gap in the time won adding more workers.

@@

\section{Performance of Heterogeneous Populations in the BBOB Benchmark}
\label{sec:exp2}

@@

@@

\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{aws}
  \caption{Boxplot of the number of evaluations per second ratio, for 15 instances of function $f_4$,
   with three execution environments Local, Single task and Multi-task.
   }
  \label{fig:aws-results}
\end{figure}

\begin{table}[]
  \small
  \caption{ Median number of the evaluations per second ratio, for 15 instances of function $f_4$,
  with three execution environments Local, Single task and Multi task}
  \label{tab:aws}
  \vspace{0.25cm}
  \centering

@@

% this is not a conclusion. The conclusion is an answer to the
% research question. This works, and works precisely for the reasons
% we say it works. - JJ
As a conclusion, EvoSwarm is a framework designed and tested according
to industry standards which is able to bring many benefits to the
bioinspired algorithm community, by showing the possibility of
creating new, high performance, mixed algorithms, which can also be
easily deployed to the cloud, or locally to systems that simply need
to have Docker and Docker Compose installed.

>>>>

  \title{EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms}
  % maybe introduce "reproducible" here, somewhere, as well as "cloud native"

@@

% Maybe these three sections should be merged; they are all experimental set up and the two next ones start with "experimental"
% (and he next one too) - JJ
% done - Mario
% But also change title? - JJ
% Experimental design of a container-based application or something like that - JJ
\section{Container-based application} 
\label{docker}

We have mentioned several advantages of a cloud-native application for running
multi-population based algorithms, but there are in our experience several
practical issues of this approach. % several issues that make it practical or useful instead? - JJ
In general, cloud platforms have free-tiers
for their services, and many offer academic discounts to students. But once we
pass to the pay-as-you-go tier, computing resources need to be
managed, and, where possible, its cost minimized.  First,
in many cases,  a credit-card is needed to guarantee the payment of the
resources consumed and someone responsible for assigning quotas and, more
importantly, special keys for using the cloud's API.  If these keys are
compromised or shared by their holders,  additional costs or administrative
procedures are in order. For those cases when a prototype platform is needed,
for instance, students developing a new algorithm variant or experimenting with
a published work, a container-based platform can be more practical. Container
technology enables researchers the local deployment of all the infrastructure
needed to run the experiments. Later, if needed, the same system could be
deployed to the cloud.

% The main advantage of container deployments over virtual machine
% instances is that they add much less overhead to communication between
% them, and they are much more economical to deploy. Scaling and
% monitoring can be done much more easily through cloud orchestration
% systems such as Kubernetes; these are the reasons why we chose it for
% this work.
% I don't see this line of argument. You can do exactly the same with
% virtual machines instead of containers. Also, you can test a virtual
% infrastructure locally and then only when it's needed deploy it to
% the cloud. If you want to give good reasons for using containers
% instead of virtual machines, there are quite a lot non-economic ones
% - JJ
% OK, I will add an issue - Mario
% Did you finally? - JJrehe
% I did not, will add and assign it - M
There are many advantages of container-based architectures over one
based on virtual machines. The main ones are architectural: it makes
the design of an application gravitate towards a set of decoupled
tasks, that use reactive, event based programming, to interact with
the rest of the application; containers hold micro-services that are
much easier to test, integrate and deploy. It also simplifies
the interaction APIs. Of course, that makes them more economical in a
pay-per-use cloud provider, but that is not the main reason why we
have chosen them in this paper; reproducibility of experiments and a
clear and idempotent description of the architecture through docker
compose % Be consistent with docker compose. Either {\tt docker-compose} or {\sf docker-compose}, but not docker compose, for instance. - JJ
are the main reasons.

In this section, we propose a design of a reactive container-based application
for executing multi-population based optimization experiments.  For the design,
we followed the design patterns highlighted in the previous section, and again
we can go back to Figure \ref{fig:kafkEO} and use it as a guide for explaining
the main components and their interactions. The containers used in this design will be explained next.

@@

% You can't start with this. Explain what this section is about, and link it with the previous one. Maybe it's better to create a big section with "Experimental setup", and make this a subsection, with subsections becoming subsubsections - JJ
Figure \ref{fig:experiment_flow} shows the general workflow to run experiments
in Evo\-Swarm, and the three roles users can play. The first step is to develop the
search strategy algorithms that are going to be used by the multi-population
algorithm; algorithm developers do this. Then administrators are responsible for
the installation and configuration of the application. Once the application is
running, users can start one or several experiments, monitor the execution, and
analyze the results. Depending on the type of experiment executed, users could
have additional needs, in this case there are additional scripts for
processing the results of experiments to be used with the COCO framework. In the
next sections, we give more details about these roles and their workflows.

@@

% You said "subsections will talk about roles", this is not a role. Change the title to that of a role. Would that be administrator? - JJ

@@

% Also, it's a bit short. Maybe merge it into the next? - JJ

% Again, not a role - JJ
\subsection{Experiment Execution}
\label{sec:exeperimentalDefinition}

@@

% This is totally unexpected. Whose role is this? You didn't even introduce it. Also, extremely short - JJ
\subsection{Using the COCO Framework}

@@

% You should clarify that, before, you were explaining how to set up an experiment. In this section, you are explaining the experiments that you are doing in this paper. But you should relate both sections to the general objectives of the paper. In the prior section, you want to prove that EvoSwarm can be used by the public to create reproductive research (and maybe you should add a small conclusion sub or sub-sub section), in the next you explain the experiment we are doing for this paper. It would be interesting to make references from the next section to the previous one, indicating how the general setup of EvoSwarm helps _us_ to create this experiment - JJ

@@

We aimed to learn if the proposed solution can efficiently
improve the scalability and performance of population-based optimization
algorithms. % If you want to prove it *improves*, you need to provide a baseline. 
            %What will be your baseline? You need to specify anyway 
            %in which way it really improves - JJ
            % The baseline is a single population GA or PSO? - M
% For instance. But I was actually looking for another word that does not imply a comparison - JJ
Hence, in the following sections, we verify the following:

@@

\begin{itemize}
\item Can we improve the execution time of the algorithm by adding populations
 and serverless functions to the system and, in particular, what is its effect on the scalability of the system? % Improve with respect to what? Single population? Please explain carefully the issues and how scaling is related to the number of populations. - JJ

\item Does having a multi-population enabled platform, with heterogeneous populations and
the support for mixing search strategies,  increase the performance of the
search by needing fewer function evaluations than a homogeneous setting? % But in this case the answer is no, right? - JJ
                                                                         % Is yes, but is in the second experiment - M
\item Can EvoSwarm as a cloud-native, container-based application,  be deployed
locally or in the cloud by just specifying the configuration of container images
and resources to an orchestration service? % This does not fit in here... It's related to reproducibility, and as such belongs in the previous section. The subsections that answer this should also go to the previous section - JJ
\end{itemize}

% Also, point forward to the section (or subsections) that will be devoted to proving each point. BTW, they should probably be subsections.

@@

% Maybe subsection? Also say the question you're answering - JJ
\section{Measuring Scalability}
\label{sec:exp1}
% Experiment 1
As we mentioned before % say where - JJ
, reactive systems can scale by adding additional copies
of serverless functions. In our case, we can start additional worker containers
to have the same effect. Other authors, like Salza et al. \cite{salza2019speed},
have used a
similar architecture, which implies that worker nodes need to have at least a certain level
of complexity, in terms of execution time,  to effectively scale on multiple
nodes, tending to linear scalability. In that paper, Salza et al. tested scalability
by simulating the work of nodes by using a sleep function and messages of different sizes.
In a population-based algorithm like EvoSwarm, several parameters can increase or 
decrease the workload of workers:

@@

To evaluate the effect of these parameters on the scalability of the system, we
propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
BBOB testbed. This function has been used because it is computationally demanding,
and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
\cite{nicolau2009application} to solve with the required FEs. As we
are only comparing in terms of execution time, the results from a single function
can be better understood, since there are fewer factors involved.
Following the procedure described in the previous section, we executed 
two sets of experiments with five and ten initial populations,
repeating each experiment 30 times using  1, 2, 4, and 8 workers.  We kept the same 
maximum number of FEs, changing the relevant parameters for this. 
See Tables \ref{tab:params:10} and \ref{tab:params:5} 
for the complete list. 
%
\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{time_diff}
  \caption{Boxplot of the median time required for the execution of 30 
  instances of function $f_4$, in seconds, different number of workers and populations sizes.  }
  % Which function? - JJ
  \label{fig:spworker}
\end{figure}
%
\begin{table}[]
  \caption{The median of the time required to finish an instance of the function $f_4$, 30 runs
  with different number of workers and populations} % time or rate? - JJ
  \label{tab:time}
  \vspace{0.25cm}
    \centering
    \begin{tabular}{ccrr}
    \hline
              &         & 5 Populations    & 10 Populations \\
     Dimension& Workers & Median of rate & Median of rate  \\
    \hline
          2   & 1w      & 2.96   & 3.15   \\
              & 2w      & 1.84   & 1.72   \\
              & 4w      & 1.22   & \textbf{0.96}   \\
              & 8w      & 1.05   & \textbf{0.59}   \\
    \hline
          3   & 1w      & 6.53   & \textbf{7.76}   \\
              & 2w      & 3.64   & 3.83   \\
              & 4w      & 2.33   & 2.26   \\
              & 8w      & 2.17   &  \textbf{1.36}   \\
    \hline
          5   & 1w      & \textbf{12.33}  & 14.19  \\
              & 2w      & 6.63   & \textbf{7.41}   \\
              & 4w      & 4.25   & 4.34   \\
              & 8w      & 3.67   &  \textbf{2.41}   \\
    \hline
          10  & 1w      & \textbf{31.50}  & 38.30  \\
              & 2w      & \textbf{16.07}  & 19.31  \\
              & 4w      &  9.99   & 10.35  \\
              & 8w      & 8.97   &  \textbf{5.92}  \\
    \hline
          20  & 1w      &  116.35 & 124.51 \\
              & 2w      & \textbf{51.18}  & 63.68  \\
              & 4w      & 32.26  & 33.07  \\
              & 8w      & 28.64  & \textbf{18.25}  \\
    \hline
          40  & 1w      & \textbf{369.89} & 448.89 \\
              & 2w      & \textbf{198.43} & 234.50 \\
              & 4w      & 122.39 &  120.62 \\
              & 8w      & 122.31 & \textbf{63.35} \\
    \hline
  \end{tabular}
\end{table}

@@

@@

% TBD? - JJ

% No idea where this fits in. I think it's the second question in 7, so it should probably be a subsection - JJ
\section{Performance of Heterogeneous Populations in the BBOB Benchmark}
\label{sec:exp2}

@@

% This should probably be a subsection of section 5, where you describe what any user needs to do to get this running and you  talk about reproducibility - JJ

@@

\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{aws}
  \caption{Boxplot of the number of evaluations per second, for 15 instances of function $f_4$,
   with three execution environments Local, Single task and Multi-task.
   }
  \label{fig:aws-results}
\end{figure}
%
\begin{table}[h!tbp]
  \small
  \caption{ Median number of the evaluations per second ratio, for 15 instances of function $f_4$,
  with three execution environments Local, Single task and Multi task}
  \label{tab:aws}
  \vspace{0.25cm}
  \centering

@@

% this is not a conclusion. The conclusion is an answer to the
% research question. This works, and works precisely for the reasons
% we say it works. - JJ
% Again, revise search questions and explain answers - JJ
As a conclusion, EvoSwarm is a framework designed and tested according
to industry standards which is able to bring many benefits to the
bioinspired algorithm community, by showing the possibility of
creating new, high performance, mixed algorithms, which can also be
easily deployed to the cloud, or locally to systems that simply need
to have Docker and Docker Compose installed.
